### 2019-12-25 report/paper.tex
report first commit

\documentclass[journal,conference]{IEEEtran}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{amsmath}
\usepackage{comment}
%\usepackage{cases}
%\usepackage{subeqnarray}
\usepackage{booktabs}
\usepackage{subfigure}
\usepackage{xcolor}
\usepackage{listings}
\lstset{
  basicstyle=\fontsize{9}{10}\selectfont\ttfamily,
  numbers=left,
  numberstyle= \tiny,
  keywordstyle= \color{ blue!70},
  commentstyle= \color{red!50!green!50!blue!50},
  frame=single,
  rulesepcolor= \color{ red!20!green!20!blue!20} ,
  escapeinside=``,
  xleftmargin=1.5em,xrightmargin=0em, aboveskip=1em,
  framexleftmargin=2em,
  showstringspaces=false,
  showtabs=false,
  breaklines=true
}
\lstdefinelanguage{Solidity}
{
  morekeywords={contract, mapping, address, uint, private, function, public, if, payable},
  morecomment=[l]{//},
  morestring=[b]"
}

\usepackage{multicol}
\usepackage{lipsum}
\usepackage{mathtools}
\usepackage{cuted}

\usepackage{amsmath}
\usepackage{extpfeil}
\usepackage{mathpartir}
\usepackage[mathscr]{eucal}

\usepackage{hyperref}
\usepackage{cleveref}

% \usepackage{CJK}

\crefformat{section}{\S#2#1#3} % see manual of cleveref, section 8.2.1
\crefformat{subsection}{\S#2#1#3}
\crefformat{subsubsection}{\S#2#1#3}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Improve SSD with Feature Fusion, \\IOU Loss and Data Augmentation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{
	Jingyu~Li,~\IEEEmembership{SJTU,}
	Haoping~Chen,~\IEEEmembership{SJTU,}
	Huangfei~Jiang,~\IEEEmembership{SJTU,}
}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~13, No.~9, September~2014}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
%
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	In this project
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
	OBI recognition, CNN, template matching, data augmentation.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\section{Introduction}
\IEEEPARstart{O}{racle}

% ljy
\section{SSD Baseline}

\section{SSD Improvement}
% ljy
\subsection{Data Augmentation}

% chp
\subsection{Feature Fusion}

% chp
\subsection{IOU Loss}

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{Ref}

% that's all folks
\end{document}